---
title: "Solution to Hands-on exercise: DAG plotting"
fontsize: 12pt
output:
  html_document:
    toc: true
    toc_depth: 2
    code_download: true
---

# Introduction

Multiple alternative exists to plot Directed Acyclic Graphs ( = DAGs = Bayesian Networks = Networks = graphs = structures). 

* abn internal functions
* Rgraphviz/DiagrammeR
* ggdag

The exact solution depends on your needs: quick visualization? paper output? interactive plotting? causal? observational analysis? fully compatible abn?

All plotting solutions requires a structure definition ... but they are different.

# ABN 

In ABN the workhorse function for plotting is *plotabn()* that uses Rgraphviz under the hood. This function Plot an ABN DAG using formula statement or a matrix.

* load abn

```{r}
library(abn)
```


* Taken that rows are children and columns are parent create a simple adjacency matrix representing the network: A -> B -> C. In using *?plotabn* what do you need additionally to plot it?

```{r}
DAG <- matrix(data = 0,nrow = 3,ncol = 3)
DAG[2,1] <- DAG[3,2] <- 1

dist<-list(A="gaussian",
           B="gaussian",
           C="gaussian")

colnames(DAG) <- rownames(DAG) <- names(dist)

```


```

* Plot the DAG

```{r}

plotabn(dag.m = DAG,data.dists = dist)

```

* Using a formula statement plot the network: A -> B -> C

```{r}

plotabn(dag.m = ~C|B+B|A,data.dists = dist)

```

* Generate an adjacency using formula statement

```{r}

ex.m <- plotabn(dag.m = ~C|B+B|A,data.dists = dist,plot = FALSE)

```


# Rgraphviz

* load the R package Rgraphviz. Using the abn function tographviz() plot the previous graph using command line.

```{r}
library("Rgraphviz")

df <- data.frame(A =1, B=1, C=1)

tographviz(dag.m = ex.m,data.df = df,data.dists = dist, outfile="graph.dot",directed=TRUE)

```

* optional*: using `https://www.bioconductor.org/packages/devel/bioc/vignettes/Rgraphviz/inst/doc/Rgraphviz.pdf` change node color.

* Using the Rpackage DiagrammeR, read the dot file created above.

```{r}
?grViz
```

# ggdag

* Load the Rpackage ggdag, plot the very same DAG as before. 

```{r}
library(ggdag)
dag  <- dagify(C~B,
               B~A)

ggdag_classic(dag, size = 6) + theme_dag()
```

* *Optional*: One strength of ggdag is to interprete causal graphs. Inspiring by the vignette (https://cran.r-project.org/web/packages/ggdag/vignettes/intro-to-dags.html) plot some causal interpretations

```{r}
dag  <- dagify(C~B,
               B~A,
labels = c("A" = "Cardiac\n Arrest", 
                  "B" = "Smoking",
                  "C" = "Cholesterol"),
       exposure = "B",
       outcome = "A")

ggdag(dag, text = TRUE, use_labels = "label")

ggdag_paths(dag, text = FALSE, use_labels = "label")
```

