---
title: "Solution to Hands-on exercise: Descriptive statistics: PID ADG"
fontsize: 12pt
output:
  html_document:
    toc: true
    toc_depth: 2
    code_download: true
---

This data set focusses on the inter-relationships among respiratory diseases (atrophic rhinitis and enzootic pneumonia), ascarid levels and daily weight gain.

# Summary statistics for the data set pig.adg

* Import the data set pig.adg

```{r}
pigs <- read.csv("data/pig_adg.csv", 
                header = TRUE, sep = ";")
```

* How many variables and how many observations are in the data set?

* Which format do the variables have? Is this correct (otherwise change)


```{r}
dim(pigs)
str(pigs)
```

```{r}
pigs$farm <- as.factor(pigs$farm)
pigs$pig <- as.factor(pigs$pig)
pigs$sex <- as.factor(pigs$sex)
pigs$pn <- as.factor(pigs$pn)
pigs$ar2 <- as.factor(pigs$ar2)
pigs$luf <- as.factor(pigs$lu)
pigs$lif <- as.factor(pigs$li)
```

```{r}
summary(pigs)
```

* Are there are any missing values?

* Table the categorical and the ordinal variables?

```{r}
table(pigs$farm)
table(pigs$sex)
table(pigs$pn)
```

* Assess by appropriate descriptive statistics (incl. also plots) if the average daily weight gain is associated with other variables in the data set?

```{r}
tapply(pigs$adg,pigs$sex,mean)
boxplot(adg ~ sex, data = pigs)

plot(adg ~ dtm, data = pigs)

plot(adg ~ mm, data = pigs)

tapply(pigs$adg,pigs$ar,mean)
boxplot(adg ~ ar, data = pigs)

tapply(pigs$adg,pigs$ar2,mean)
boxplot(adg ~ ar2, data = pigs)

tapply(pigs$adg,pigs$lu,mean)
boxplot(adg ~ lu, data = pigs)

tapply(pigs$adg,pigs$pn,mean)
boxplot(adg ~ pn, data = pigs)

plot(adg ~ epg5, data = pigs)

plot(adg ~ worms, data = pigs)

tapply(pigs$adg,pigs$li,mean)
boxplot(adg ~ li, data = pigs)
```

* Are the scores for the atrophic rhinitis and the enzootic pneumonia possibly associated?

```{r}
table(pigs$ar,pigs$lu)
table(pigs$ar2,pigs$pn)
```

```{r}
mosaicplot(~ pn + ar2 + li, data = pigs)
```
