---
title: "Solution to Hands-on exercise: Simulation of an ABN model"
fontsize: 12pt
output:
  html_document:
    toc: true
    toc_depth: 2
    code_download: true
---

# Manually simulate abn data

* Generate a dataframe of 1000 observations, based on the following set of statements (purely gaussian network): Z influences X (0.7), z influences y (0.5), x influnences y (1). All nodes have a random error. 

```{r}
set.seed(42)

n <- 1000
z <- rnorm(n = n,mean = 0,sd = 0.1)
x <- 0.7*z + rnorm(n = n,mean = 0,sd = 0.1)
y <- 1*x + 0.5*z + rnorm(n = n,mean = 0,sd = 0.1)

df <- data.frame(x,y,z)

```

* Using buildscorecache() and mostprobable() retrieve the simulated network

```{r}
dist <- list(x = "gaussian",y = "gaussian",z = "gaussian")

bsc <- buildscorecache(data.df = df,data.dists = dist,max.parents = 2)

dag <- mostprobable(score.cache = bsc)

plotabn(dag.m = (dag),data.dists = dist,edgedir = "cp")

```



# Simulate data using abn functions

* Using simulateabn() simulate dataset from the given DAG. You need to define a number of observations, a matrix of parameters and a matrix of variance

```{r}
dist <- list(a="gaussian",
b="gaussian",
c="gaussian",
d="gaussian", 
e="binomial", 
f="binomial")

plotabn(dag.m = ~a|b:c:e+b|c:d:f+e|f,data.dists = dist ,plot = TRUE)

```

```{r}
#define parameter matrix
data.param <- matrix(data = c(1,2,0.5,0,20,0,
                              0,1,3,10,0, 0.8,
                              0,0,1,0,0,0,
                              0,0,0,1,0,0,
                              0,0,0,0,0.5,1,
                              0,0,0,0,0,0),nrow = 6L,ncol = 6L,byrow = TRUE)

#precision matrix 
data.param.var <- matrix(data = 0,nrow = 6L,ncol = 6L)
diag(data.param.var) <- c(10,20,30,40,0,0)

#simulate the data
df <- simulateabn(data.dists = dist,
n.chains = 1,
n.thin = 1,
n.iter = 1000,
data.param = data.param,
data.param.var = data.param.var,
seed = 132)

```

* Based on the simulated dataset perform a BN search using buildscorecache() and mostprobable()

```{r}
bsc <- buildscorecache(data.df = df,data.dists = dist,max.parents = 2)

dag <- mostprobable(score.cache = bsc)

plotabn(dag.m = (dag),data.dists = dist,edgedir = "cp")

```

* Optional: perform a heuristic search.

```{r}
bsc <- buildscorecache(data.df = df,data.dists = dist,max.parents = 2)

dag <- search.heuristic(score.cache = bsc,score = "mlik",data.dists = dist,max.parents = 1,num.searches = 500,max.steps = 100)

plotabn(dag.m = (dag$dag),data.dists = dist,edgedir = "cp")

```
